---
title: "R Notebook"
output: html_notebook
---

```{r}
# importing the data and reading the top 10 records using head() function
cred_data <- read.csv(file.choose(), stringsAsFactors = TRUE)
head(cred_data,10)
str(cred_data)
```

```{r}
# Checking the summary of data an finding histogram of target variable. Most people are having account balance below 500.
summary(cred_data)
hist(cred_data$Balance)
```


# Question 1: Linear Regression of Balance with individual explanatory variables.

```{r}
# Linear Regression: Balance vs Income
lr_incomeBal <- lm(Balance~Income, data = cred_data)
summary(lr_incomeBal)
plot(Balance~Income,data = cred_data, main = 'Balance vs Income')
abline(lr_incomeBal)
```
```{r}
# Linear Regression: Balance vs Limit
lr_limitBal <- lm(Balance~Limit, data = cred_data)
summary(lr_limitBal)
plot(Balance~Limit,data = cred_data, main = 'Balance vs Limit')
abline(lr_limitBal)
```




```{r}
# Linear Regression: Balance vs CreditRating
lr_credratBal <- lm(Balance~CreditRating, data = cred_data)
summary(lr_credratBal)
plot(Balance~CreditRating,data = cred_data, main = 'Balance vs CreditRating')
abline(lr_credratBal)
```


```{r}
# Linear Regression: Balance vs Age
lr_ageBal <- lm(Balance~Age, data = cred_data)
summary(lr_ageBal)
plot(Balance~Age,data = cred_data, main = 'Balance vs Age')
abline(lr_ageBal)
```


```{r}
# Linear Regression: Balance vs Eduation
lr_eduBal <- lm(Balance~Education, data = cred_data)
summary(lr_eduBal)
plot(Balance~Education,data = cred_data, main = 'Balance vs Education')
abline(lr_eduBal)
```



```{r}
# Linear Regression: Balance vs Student
# converting categorical variable Student to numeric variable
cred_data$Student<-factor(as.numeric(cred_data$Student))

# performing linear regression
lr_studentBal <- lm(Balance~Student, data = cred_data)
summary(lr_studentBal)
plot(Balance~Student,data = cred_data, main = 'Balance vs Student')
abline(lr_studentBal)
```


```{r}
# Linear Regression: Balance vs Married
# converting categorical variable Student to numeric variable
cred_data$Married<-factor(as.numeric(cred_data$Married))

# performing linear regression
lr_marriedBal <- lm(Balance~Married, data = cred_data)
summary(lr_marriedBal)
plot(Balance~Married,data = cred_data, main = 'Balance vs Married')
abline(lr_marriedBal)
```

Ranking the covariates by the highest R2 order

Credit Rating: 0.7458484

Limit: 0.7425222

Income: 0.2149773

Student: 0.06709009

Married: 3.218849e-05

Education: 6.498901e-05

Age: 3.367661e-06

From the above regression models and their R2 values, it is obtained that Credit Rating and Limit are having
the highest R2 value.


# Question 2: Full Model Regression

```{r}
# Regression with full model including all variables as explanatory variables
lr_full <- lm(Balance ~ ., data = cred_data)
summary(lr_full)
```


Here we can see that RSE is very low as compared to all linear models obtained by regressing individual variables with dependent variables.

Also, Looking a p-values, it seems like Income, Limit, CreditRating, Student and somewhat Age are significant variables indeciding how well the value of Balance can be predicted.

We can verify this by building a model excluding Education and Married.

```{r}
# Linear model to verify the significant variables
lr_onlysignificant <- lm(Balance ~ Income + Limit + Student + CreditRating + Age, data = cred_data)
summary(lr_onlysignificant)
```

From above result, we can see that RSE has not changed a lot by removing Education and Married. This verifies that rest of the variables are the only significant ones.




# Question 3: Deciding whether CreditRating is statistically significant or not for a perfectly fitting linear model.

For this, we will construct a linear model without Credit Rating and compare its coefficients with full model

```{r}
# Regression with full model excluding  CreditRating.
lr_nocred <- lm(Balance ~ Income + Limit + Student + Age + Education + Married, data = cred_data)
summary(lr_nocred)
```


Comparing the two models: lr_full and lr_nocred,

            lr_full       lr_nocred
RSE           100.8           104.2
DOF           392             393
Median      -12.67          -18.68
Adj. R2       0.952           0.948


lr_full is a linear model with CreditRating included as a regressor
lr_nocred is a linear model with CreditRating not included as a regressor

It can be seen from above comparision that RSE has increased by ~4, Ajusted R2 has decreased by 0.004 and median has moved a little bit away from zero by 6.

This means the model is not the best fit without credit rating but it is not having a major impact on the results. Even without third party credit rating score, internal bank data can efficiently predict the balance of a new coming customer. The models can also be further optimized by applying regularization methods. 

# Business recommendation: 
If the credit scoring system is really expensive for the business, it can be excluded from the analysis given the customer base is not very wide and pose a large amount of financial risk for the bank if special interest rates are accidentally given to a customer. In this case, a cost analysis is necessary to determine if savings from not taking the credit rating score will be more or from not ending up giving special interest rates to unpromising customers.
















